---
title: "SDM"
author: "Lallu Nikerthil Prathapan"
date: "4/18/2022"
output: html_document
---

```{r knitr_init, echo=FALSE, cache=FALSE}
library(knitr)
install.packages("rmdformats")
library(rmdformats)
## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

#Installing the necessary packages for running the SDM

```{r Package installation, eval=FALSE, echo=FALSE}
my_packages<-c("rgdal","raster","rgeos","dismo","letsR","biomod2","biogeo","maptools","stringr","ggplot2","plotly","xtable","tidyr","rlang")
lapply(my_packages,require,character.only=TRUE)#loads multiple packages
```

```{r Loading libraries, echo=FALSE}
library(rgdal)
library(raster)
library(rgeos)
library(dismo)
library(letsR)
library(biomod2)
library(biogeo)
library(maptools)
library(stringr)
library(ggplot2)
library(plotly)
library(xtable)
library(tidyr)
```

#* This project is build,evaluate and project species distribution model for Panthera species in South American region(Neotropical zoogeographical region ) as the study area.

*Load study area*
SouthAmerica_shape <- rgdal::readOGR("D:/EAGLE/MB2/SDM_tutorial_v3/data/shapefiles/neotropical.shp")
plot(SouthAmerica_shape)