---
title: "SDM"
author: "Lallu Nikerthil Prathapan"
date: "4/18/2022"
email:"lalsdestiny@gmail.com"
output: html_document
---
#This project is build,evaluate and project species distribution model in South American region(Neotropical zoogeographical region ) as the study area.

## The step involved in the process are below
###1.Data preparation
#     +Study area preparation
#     +Environmental data import
#     +Predictor variable extraction
#     +Presence and absence data
#     +Presence-only and pseudo-absences data
###2.Model fitting,prediction and evaluation
###3.Projecting models
```{r knitr_init, echo=FALSE, cache=FALSE}
library(knitr)
install.packages("rmdformats")
library(rmdformats)
## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

#Installing the necessary packages for running the SDM

```{r Package installation, eval=FALSE, echo=FALSE}
my_packages<-c("rgdal","raster","rgeos","dismo","letsR","biomod2","biogeo","maptools","stringr","ggplot2","plotly","xtable","tidyr","rlang")
lapply(my_packages,require,character.only=TRUE)#loads multiple packages
```

```{r Loading libraries, echo=FALSE}
library(rgdal)
library(raster)
library(rgeos)
library(dismo)
library(letsR)
library(biomod2)
library(biogeo)
library(maptools)
library(stringr)
library(ggplot2)
library(plotly)
library(xtable)
library(tidyr)
```
#Step1-Data preparation

*Load study area and creating spatial grids for study area*
```{r Loading study area shapefile, echo = TRUE}
SouthAmerica_shape <- rgdal::readOGR("D:/EAGLE/MB2/SDM/SDM_MB2/data/shapefiles/neotropical.shp")
plot(SouthAmerica_shape)
```
#Defines a function for creating study area grid with resolution and proportion of overlap for each cell of the grid
```{r GridFilter function, echo = TRUE}
grid_filter <- function(shape, resol = 1, prop = 0){
  grid <- raster(extent(shape))
  res(grid) <- resol
  proj4string(grid) <- proj4string(shape)
  polygon_grid <- rasterToPolygons(grid)
  proj_dryland <- spTransform(shape, CRS("+proj=laea"))
  proj_grid_poly <- spTransform(polygon_grid, CRS("+proj=laea"))
  proj_grid_poly$layer <- c(1:length(proj_grid_poly$layer))
  area_grid <- gArea(proj_grid_poly, byid=T)
  proj_grid_poly <- gBuffer(proj_grid_poly, byid=TRUE, width=0)
  land.grid <- intersect( proj_dryland, proj_grid_poly)
  land.grid_area <- gArea(land.grid, byid=T)
  info <- cbind(land.grid$layer, area_grid[land.grid$layer],     land.grid_area)
  land.grid$layer <- info[,3]/info[,2]
  land.grid <- spTransform(land.grid, CRS(proj4string(shape)))
  land.grid.filtered <- land.grid[land.grid$layer >= prop,]
}
```

# A spatial polygon gird of resolution 5degreex5degree is created
```{r polygond grid,echo=FALSE}
SA_grid <- grid_filter(SouthAmerica_shape, resol = 5, prop = 0.5)
#ID column is added for future use
SA_grid$ID <- 1:(length(SA_grid))
#Exporting the grid to a file
writeOGR(SA_grid, dsn=paste("D:/EAGLE/MB2/SDM/SDM_MB2/data/shapefiles", sep=""), layer="South_America_grid", driver="ESRI Shapefile", overwrite_layer=T)
```

#Plotting resulting polygon-each cell is a site(a row) in the data
```{r Loading and plotting your polygon grid, echo=TRUE, warning=FALSE}
SA_grid <-  readOGR("D:/EAGLE/MB2/SDM/SDM_MB2/data/shapefiles/South_America_grid.shp")
plot(SA_grid)
```
##Step2- Environmental data and extraction of predictor variable

`The predictor variables for SDM are extracted from the raster file of environmental variables like climate, soil, terrain, landuse etc. Here the climate variable is taken into consideration `

*Importing environmental data*

```{r Improting bioclimatic variables locally, echo = TRUE}
#Using locally downloaded data as predictor variable
clim_rasters<- list.files("D:/EAGLE/MB2/SDM/SDM_MB2/data/rasters/bio_10m_bil", pattern = "bil$", full.names = TRUE)
clim_rasters

#Stacking the Climate variables
clim_rasters<-stack(clim_rasters)
clim_rasters

#Plotting the climate variables
plot(clim_rasters/10) #Worlclim data is multiplied by 10.
```

```{r Bio climate table definition, tidy=FALSE, echo = FALSE}
kable(
  data.frame(
    var= c("BIO1","BIO2","BIO3","BIO4","BIO5","BIO6","BIO7","BIO8","BIO9","BIO10","BIO11","BIO12","BIO13","BIO14","BIO15","BIO16","BIO17","BIO18","BIO19"),
    Definition = c(" Annual Mean Temperature","Mean Diurnal Range (Mean of monthly (max temp - min temp))","Isothermality (BIO2/BIO7) (×100)","Temperature Seasonality (standard deviation ×100)","Max Temperature of Warmest Month","Min Temperature of Coldest Month","Temperature Annual Range (BIO5-BIO6)","Mean Temperature of Wettest Quarter","Mean Temperature of Driest Quarter","Mean Temperature of Warmest Quarter"," Mean Temperature of Coldest Quarter","Annual Precipitation","Precipitation of Wettest Month","Precipitation of Driest Month"," Precipitation Seasonality (Coefficient of Variation)","Precipitation of Wettest Quarter"," Precipitation of Driest Quarter","Precipitation of Warmest Quarter","Precipitation of Coldest Quarter")
            ),
  caption = "Bioclimatic variables are derived from the monthly temperature and rainfall values in order to generate more biologically meaningful variables. These are often used in species distribution modeling and related ecological modeling techniques. The bioclimatic variables represent annual trends (e.g., mean annual temperature, annual precipitation) seasonality (e.g., annual range in temperature and precipitation) and extreme or limiting environmental factors (e.g., temperature of the coldest and warmest month, and precipitation of the wet and dry quarters). A quarter is a period of three months (1/4 of the year)."#, booktabs = TRUE
)
```

*Converting the climate rasters and shapefile to same projection for extraction*
```{r Defining projections, echo = TRUE}
proj.coord <- CRS("+proj=longlat +ellps=WGS84 + datum=WGS84")  # geographical, datum WGS84
SA_grid <- spTransform(SA_grid, proj.coord)  # projectind the study area grid
proj4string(clim_rasters) <- SA_grid  # and of climatic rasters
```

#Here for the species panthera the  bioclimate variable 1,2,3,4,8,12,15,18 are considered as the predictor variables
*Extraction of predictor variable for each cell of the grid*
```{r Variable extraction 1, echo = TRUE}
expl_var<-data.frame(bio_1=numeric(length(SA_grid)),bio_2=numeric(length(SA_grid)),bio_3=numeric(length(SA_grid)),
                     bio_4=numeric(length(SA_grid)),bio_8=numeric(length(SA_grid)),bio_12=numeric(length(SA_grid)),
                     bio_15=numeric(length(SA_grid)),bio_18=numeric(length(SA_grid)))
expl_var

#extr_1 is a list of cells that contains all each cell for each predictor
extr_1<-raster::extract(clim_rasters$bio1,SA_grid)
extr_2 <- raster::extract(clim_rasters$bio2,SA_grid)
extr_3 <- raster::extract(clim_rasters$bio3,SA_grid)
extr_4 <- raster::extract(clim_rasters$bio4,SA_grid)
extr_8 <- raster::extract(clim_rasters$bio8,SA_grid)
extr_12 <- raster::extract(clim_rasters$bio12,SA_grid)
extr_15 <- raster::extract(clim_rasters$bio15,SA_grid)
extr_18 <- raster::extract(clim_rasters$bio18,SA_grid)

expl_var$bio_1<-unlist(lapply(extr_1,function(x) 
                                if(!is.null(x)) 
                                  mean(x,na.rm=TRUE)
                                else NA))
expl_var$bio_2<-unlist(lapply(extr_2,function(x) 
                         if(!is.null(x)) 
                           mean(x,na.rm=TRUE)
                         else NA))  
expl_var$bio_3<-unlist(lapply(extr_3,function(x) 
                         if(!is.null(x)) 
                           mean(x,na.rm=TRUE)
                         else NA))
expl_var$bio_4<-unlist(lapply(extr_4,function(x) 
                         if(!is.null(x)) 
                           mean(x,na.rm=TRUE)
                         else NA))
expl_var$bio_8<-unlist(lapply(extr_8,function(x) 
                         if(!is.null(x)) 
                           mean(x,na.rm=TRUE)
                         else NA))
expl_var$bio_12<-unlist(lapply(extr_12,function(x) 
                         if(!is.null(x)) 
                           mean(x,na.rm=TRUE)
                         else NA))
expl_var$bio_15<-unlist(lapply(extr_15,function(x) 
                          if(!is.null(x)) 
                            mean(x,na.rm=TRUE)
                          else NA))
expl_var$bio_18<-unlist(lapply(extr_18,function(x) 
                          if(!is.null(x)) 
                            mean(x,na.rm=TRUE)
                          else NA))
write.table(expl_var,"D:/EAGLE/MB2/SDM/SDM_MB2/data/matrices/Environ_var_exracted.txt")
head(expl_var)

``` 
## Importing species data
### Using expert drawn maps

Here the species distribution information is acquired from the IUCN Red List for Threatened Species database. These expert drawn maps acquired by multiple method:only on occurence data(e.g. polygon around known occurence)

One way of acquiring species distribution information is from expert-drawn maps, such as the ones from the IUCN Red List for Threatened Species database. These maps Tare usually done used multiple methods: only on occurrence data (e.g. a polygon around known occurrences), whereas others incorporate multiple degrees of knowledge, such as habitat requirements or elevational limits of the species, in essence using an informal distribution modelling approach. 

Richness maps derived from expert-drawn range maps may overestimate local richness ('error of commission'), in relation to point-to-grid richness maps. This because they are generally drawn to include all areas where a species is known to occur without excluding areas in between where the species may not exist. They tend to map the 'extent of occurrence' of species that includes the, perhaps much smaller, ‘area of occupancy’ (Loislle et al., 2003; Habib et al., 2004; Hurlbert & White, 2005; Graham & Hijmans, 2006). 

```{r Plotting IUCN distributions, echo=TRUE, warning=FALSE}
#Loading Species map file
species_data<-readShapePoly("D:/EAGLE/MB2/SDM/SDM_MB2/data/shapefiles/NT_TERRESTRIAL_MAMMALS_subset.shp",proj4string = proj.coord)
plot(species_data)
```

*Creates presence-absence matrix of all species within the study area grid using* `lets.presab.grid` *from the* `letsR` *package.*

```{r presence-absence matrix creation, echo = TRUE}
proj.coord=CRS("+proj=longlat +ellps=WGS84 + datum=WGS84") #geographical, datum
proj4string(SA_grid)<-proj.coord  #projecting the grid data
proj4string(species_data)<-proj.coord

pres_abs.matrix<-lets.presab.grid(shapes = species_data,
                                  grid=SA_grid,
                                  sample.unit = "ID",
                                  remove.sp = TRUE,
                                  presence = NULL,
                                  origin=NULL,
                                  seasonal = NULL)

#Exporting the presence absence matrix
write.table(pres_abs.matrix$PAM,"D:/EAGLE/MB2/SDM/SDM_MB2/data/matrices/Presence_absence_matrix.txt")

```

*Plot the species richness within the study grid

```{r Richness map, echo = TRUE}
rich_plus<-rowSums(pres_abs.matrix$PAM)+1
colfunc<-colorRampPalette(c("#fff5f0", "#fb6a4a", "#67000d"))
colors<- c("white",colfunc(max(rich_plus)))
plot(pres_abs.matrix$grid,border="gray40",col=colors[rich_plus])
map(add=TRUE)
```
#Extracting the Coordinates of cell centeroids for further data analysis
```{r Extracting cell centroids coordinates, echo=TRUE}
resp_coord_var<-as.data.frame(coordinates(pres_abs.matrix$grid))
colnames(resp_coord_var)<-c("Longitude X", "Latitude Y")

#Export to file
write.table(resp_coord_var,"D:/EAGLE/MB2/SDM/SDM_MB2/data/matrices/coordinates_centeroids.txt")
```

#Data cleaning
```{r data cleaning, echo=TRUE}
#Species data
species<-read.table("D:/EAGLE/MB2/SDM/SDM_MB2/data/matrices/Presence_absence_matrix.txt",header=TRUE)
head(species)
colSums(species)

#removing the columns having less than 4 presence
rem_col<-colnames(species)[colSums(species)<=4]
`%ni%`<-Negate(`%in%`)
species<-subset(species,select=names(species) %ni% rem_col)

#change the separter between species name using 'gsub'function
#Replace "." per "."
names(species)<-gsub(x=names(species),pattern="\\.",replacement = ".")
write.table(species,"D:/EAGLE/MB2/SDM/SDM_MB2/data/matrices/Presence_absence_matrix_cleaned.txt")

```

#Species distribution modelling

*Arranging the input data for usage of the function `BIOMOD_FormatingData()` 

```{r format data, echo=TRUE}
#Load species data
resp_var_data<-read.table("D:/EAGLE/MB2/SDM/SDM_MB2/data/matrices/Presence_absence_matrix_cleaned.txt",header=TRUE)
#Load environmental variables
expl_var_data<-read.table("D:/EAGLE/MB2/SDM/SDM_MB2/data/matrices/Environ_var_exracted.txt")
#Define species of Interest-selecting Panthera.onca
species_name<-names(resp_var_data)[1]
species_name
#Loading coordinates
resp_coord_var<-read.table("D:/EAGLE/MB2/SDM/SDM_MB2/data/matrices/coordinates_centeroids.txt")
```

```{r BIOMOD_FormatingData,echo=TRUE}
modelling_data<-BIOMOD_FormatingData(resp.var =as.data.frame(resp_var_data[,species_name]),
                                     expl.var =expl_var_data,
                                     resp.xy = as.data.frame(resp_coord_var),
                                     resp.name = species_name,
                                     #PA.nb.rep = 2,
                                     #PA.nb.absences = 200,
                                     #PA.strategy = 'random',
                                     na.rm=TRUE)
modelling_data
```
#Plot the data
```{r plot BIOMOD_FormatingData,echo=TRUE}
plot(modelling_data)
legend(x="bottomleft",
       legend=c("absent","present"),
       fill=2:3,
             cex=1,
       title=NULL)

```

#run models
#Setting default parameters for the modelling techniques using function BIOMOD_ModelingOptions()
```{r BIOMOD_ModelingOptions,echo=TRUE}
modelling_option<-BIOMOD_ModelingOptions()
```

##Computing models
#Here a set of models will be run for the panthera species. Selected algorithms are 'GLM','GBM','RF'
## As there is no evaluation data 3 fold cross validation is done  by randomly splitting the data set to 2 subsets
```{r Biomodeloutput,echo=TRUE}
model_output<-BIOMOD_Modeling(
#BIOMOD_FormatingData()-class object
data = modelling_data,
#Model sets to be calibrated on data
models=c('GBM','GLM','RF'),
#Number of evaluation to run
NbRunEval = 3,
#70% data will be run and rest is used for testing
DataSplit = 70,
#Prevelance when kept null, presence-absence have same weight
Prevalence = 0.5,
#Number of permutations to estimate variable importance
VarImport = 5,
#Evaluation methods
models.eval.meth = c('TSS','ROC'),
#save results on harddrive
SaveObj = TRUE,
#when the afrgument TRUE, model prediction have binomial scale
rescal.all.models = FALSE,
#when the argument TRUE, model is calibrated and evaluated with whole dataset
do.full.models = FALSE,
#An Id for modelling procedure
modeling.id = paste(species_name,"Climate",sep = ""))

```
##Model Evaluation
##Extracting evaluation scores
#Receiving Operator Curve(ROC) and True Skill Statistic(TSS) evaluations are used 
```{r model evaluation, echo=TRUE}
#get all model evaluation
model_evalution<-get_evaluations(model_output)#evaluation scores 
model_evalution

#print the dimnames of the object
dimnames(model_evalution)
#histogram of model evalution
x11()
hist(model_evalution)
write.table(model_evalution,file="D:/EAGLE/MB2/SDM/SDM_MB2/data/results/panthera_model_evaluation")

```

#Plotting evaluation scores-Sensitivity and Specificity

#TSS and ROC score for sensitivity
```{r Sensitivity, echo=TRUE,eval=FALSE}
model_sensitivity<-model_evalution[,"Sensitivity",,,]
model_sensitivity

all_model_sens<-as.data.frame(t(as.data.frame(model_sensitivity)))

sens.hist.TSS<-ggplot(all_model_sens,aes(x=TSS))+geom_histogram(aes(y=..density..),color="black",fill='grey',
                      binwidth = density(all_model_sens$TSS)$bw)+geom_density(fill='red',alpha=0.2)
                     

sens.hist.ROC<-ggplot(all_model_sens,aes(x= ROC))+geom_histogram(aes(y=..density..),color="black",fill='grey',
                                                                binwidth = density(all_model_sens$ROC)$bw)+geom_density(fill='red',alpha=0.2)
sens.hist.TSS
sens.hist.ROC

write.table(model_sensitivity,file="D:/EAGLE/MB2/SDM/SDM_MB2/data/results/sensitivity_score_3models")

```
#TSS and ROC score for Specificity

#SPECIFICITY OF ALL MODELS-TSS AND ROC SCORE FOR SPECIFICITY


```{r}
#SPECIFICITY OF ALL MODELS-TSS AND ROC SCORE FOR SPECIFICITY
model_specificity<-model_evalution[,"Specificity",,,]
model_specificity
hist(model_specificity)

write.table(model_specificity,file = "D:/EAGLE/MB2/SDM/SDM_MB2/data/results/specificity_score_3models")
```
#PRESENCES proportion correctly predicted (true positive)
```{r true positive,}
sensitivity<-read.table("D:/EAGLE/MB2/SDM/SDM_MB2/data/results/sensitivity_score_3models")
summary(sensitivity)
str(sensitivity)
sensitivity #table of sensitivity by TSS or ROC run per model
```
# plotting box plot for diff algorithm
#TSS score
```{r sensitivity_TSS_plot, echo=TRUE}
TSS.sens<-sensitivity[1,]
trans.sens<-t(TSS.sens)#Matrix transposing for getting table
modelName<-rownames(trans.sens)
model1<-substr(modelName,1,3)
sens_plot.1<-data.frame(trans.sens,model1)
```

```{r sensitivity_ROC_plot, echo=TRUE}
#ROC score
ROC.sens<-sensitivity[2,]
trans.sens<-t(ROC.sens)#Matrix transposing for getting table
modelName<-rownames(trans.sens)
model2<-substr(modelName,1,3)
sens_plot.2<-data.frame(trans.sens,model2)
```

```{r}
sens_plot_data<-cbind(sens_plot.1,sens_plot.2)
```

#Box plot of sensitivity score for ROC and TSS
```{r  sensitivity box plot ,echo=TRUE}
TSS.sens.bplot<-ggplot(sens_plot_data,aes(x=model1,y=TSS,fill=model1))+
                geom_boxplot()+
                guides(fill=FALSE)+
                ggtitle("Sensitivity using TSS scores")+
                theme(plot.subtitle = element_text(vjust = 1), 
                plot.caption = element_text(vjust = 1), 
                panel.background = element_rect(fill = NA)) +
                labs(x = "Model Type", y = "TSS (%)") 
TSS.sens.bplot<-ggplotly(TSS.sens.bplot)

ROC.sens.bplot<-ggplot(sens_plot_data,aes(x=model2,y=ROC,fill=model2))+
                geom_boxplot()+
                guides(fill=FALSE)+
                ggtitle("Sensitivity using ROC scores")+
                theme(plot.subtitle = element_text(vjust = 1), 
                plot.caption = element_text(vjust = 1), 
                panel.background = element_rect(fill = NA)) +
                labs(x = "Model Type", y = "ROC (%)") 
ROC.sens.bplot<-ggplotly(ROC.sens.bplot)

TSS.sens.bplot
ROC.sens.bplot
```
#ABSENCE proportion correctly predicted (true negative)

```{r true negative, echo=TRUE}
specificity<-read.table("D:/EAGLE/MB2/SDM/SDM_MB2/data/results/specificity_score_3models")
summary(specificity)
str(specificity)
specificity#table of specificity by TSS or ROC run per model
```
# plotting box plot for diff algorithm
#TSS score
```{r specificity_TSS_plot, echo=TRUE}
TSS.speci<-specificity[1,]
trans.speci<-t(TSS.speci)#Matrix transposing for getting table
modelName<-rownames(trans.speci)
model1<-substr(modelName,1,3)
speci_plot.1<-data.frame(trans.speci,model1)
```

```{r specificity_ROC_plot, echo=TRUE}
#ROC scor
ROC.speci<-specificity[2,]
trans.speci<-t(ROC.speci)#Matrix transposing for getting table
modelName<-rownames(trans.speci)
model2<-substr(modelName,1,3)
speci_plot.2<-data.frame(trans.speci,model2)
```

```{r}
speci_plot_data<-cbind(speci_plot.1,speci_plot.2)
```

#Box plot of sensitivity score for ROC and TSS
```{r  specificity box plot ,echo=TRUE}
TSS.speci.bplot<-ggplot(speci_plot_data,aes(x=model1,y=TSS,fill=model1))+
  geom_boxplot()+
  guides(fill=FALSE)+
  ggtitle("Specivicity using TSS scores")+
  theme(plot.subtitle = element_text(vjust = 1), 
        plot.caption = element_text(vjust = 1), 
        panel.background = element_rect(fill = NA)) +
  labs(x = "Model Type", y = "TSS (%)") 
TSS.speci.bplot<-ggplotly(TSS.speci.bplot)

ROC.speci.bplot<-ggplot(speci_plot_data,aes(x=model2,y=ROC,fill=model2))+
  geom_boxplot()+
  guides(fill=FALSE)+
  ggtitle("Specivicity using ROC scores")+
  theme(plot.subtitle = element_text(vjust = 1), 
        plot.caption = element_text(vjust = 1), 
        panel.background = element_rect(fill = NA)) +
  labs(x = "Model Type", y = "ROC (%)") 
ROC.speci.bplot<-ggplotly(ROC.speci.bplot)

TSS.speci.bplot
ROC.speci.bplot
```

#Plotting variable importance and saving it
```{r variable importance,echo=TRUE}
get_variables_importance(model_output)
#saving the variables importance and model evaluation
capture.output(get_evaluations(model_output),file = file.path(species_name,paste(species_name,"model evaluation.txt",sep="")))
capture.output(get_variables_importance(model_output),file = file.path(species_name,paste(species_name,"model_variable_importance.txt",sep="")))
```
#Plotting the response curves for one run
```{r 2D response curve GLM GBM RF}
#GLM
GLM<-BIOMOD_LoadModels(model_output,models=c("GLM"))
#2D response curve
Resplot2D_GLM <- response.plot2(models  = GLM,
                               Data = get_formal_data(model_output,'expl.var'), 
                               show.variables= get_formal_data(model_output,'expl.var.names'),
                               do.bivariate = FALSE,
                               fixed.var.metric = 'median',
                               col = c("blue", "red"),
                               legend = TRUE,
                               data_species = get_formal_data(model_output,'resp.var'))
#GAM
GBM<-BIOMOD_LoadModels(model_output,models=c("GBM"))
#2D response curve
Resplot2D_GBM <- response.plot2(models  = GBM,
                            Data = get_formal_data(model_output,'expl.var'), 
                            show.variables= get_formal_data(model_output,'expl.var.names'),
                            do.bivariate = FALSE,
                            fixed.var.metric = 'median',
                            col = c("blue", "red"),
                            legend = TRUE,
                            data_species = get_formal_data(model_output,'resp.var'))
#RF
RF<-BIOMOD_LoadModels(model_output,models=c("RF"))
#2D response curve
Resplot2D_RF <- response.plot2(models  = RF,
                                Data = get_formal_data(model_output,'expl.var'), 
                                show.variables= get_formal_data(model_output,'expl.var.names'),
                                do.bivariate = FALSE,
                                fixed.var.metric = 'median',
                                col = c("blue", "red"),
                                legend = TRUE,
                                data_species = get_formal_data(model_output,'resp.var'))
```
#3D RESPONSE PLOT for one model run
```{r  2D response curve GLM GBM RF}
#GLM
x11()
Resplot3D_GLM <- response.plot2(models  = GLM[1],
                               Data = get_formal_data(model_output,'expl.var'), 
                               show.variables= get_formal_data(model_output,'expl.var.names'),
                               do.bivariate = TRUE,
                               fixed.var.metric = 'median',
                               data_species = get_formal_data(model_output,'resp.var'),
                               display_title=FALSE)
#GBM
X11()
Resplot3D_GBM <- response.plot2(models  = GBM[1],
                                Data = get_formal_data(model_output,'expl.var'), 
                                show.variables= get_formal_data(model_output,'expl.var.names'),
                                do.bivariate = TRUE,
                                fixed.var.metric = 'median',
                                data_species = get_formal_data(model_output,'resp.var'),
                                display_title=FALSE)
#RF
X11()
Resplot3D_RF <- response.plot2(models  = RF[1],
                                Data = get_formal_data(model_output,'expl.var'), 
                                show.variables= get_formal_data(model_output,'expl.var.names'),
                                do.bivariate = TRUE,
                                fixed.var.metric = 'median',
                                data_species = get_formal_data(model_output,'resp.var'),
                                display_title=FALSE)

```

#PLOTTING PREDICTIONS
#Make projections on the current variable

```{r Plotting predictions, echo=TRUE}
biomodproj <- BIOMOD_Projection(
  modeling.output = model_output,
  new.env = expl_var_data,
  xy.new.env = resp_coord_var,
  proj.name = 'current',
  selected.models = 'all',
  binary.meth = 'ROC',
  compress = TRUE,
  clamping.mask = F,
  output.format = '.RData')

# Plot predictions
plot(biomodproj)
SDM_project <- get_predictions(biomodproj)
SDM_project
```






